Integer Programming from Quantum Annealing and Open Quantum Systems

### 2020-09-23 paper/main.tex
Updated equations to match main section and cross-checked with code

The solution to the Minimum Dominating Set problem can be expressed as an integer optimization problem given by,
\begin{align}
 f(\vec x) = & \min\left(\sum_{v \in V} x_v\right),                    \\
\end{align}
subject to
\begin{align}
 & x_v + \sum_{j \in \mathcal{N}(v)} x_j \geq 1, \\
 & x_v \in \{0, 1\}
\end{align}
where $\mathcal{N}(v)$ represents the set of all direct neighbors of vertex $v$ and the dimension of the dependent variable $x$ is the number of vertices $n_V \equiv \overline{\overline{V}}$.
The problem minimizes the number of vertices in $D$ with a binary variable $\vec x = \vec \Psi_x$ encoded by single qubit, subject to the constraint that at least one vertex in $\mathcal{N}(\nu)$ is in $D$.
For each vertex in $V$ we introduce slack variables
\begin{equation}
    \vec s \in \left\{ \mathbb{Z}^{n_V} \, \middle| \, 0 \leq s_{v} \leq |\mathcal{N}(v)| \quad \forall v\in V \right\} \, ,
\end{equation}
which is related to the qubit vector $\vec \Psi_s$ by $\vec s = T_s \vec \Psi_s$, to encode the inequality constraint such that
\begin{align}
 f(\vec \Psi_x)
 =
 & \min\left(\vec 1 \cdot \vec \Psi_x \right) \, ,
\end{align}
subject to
\begin{align}
 &
 \vec \Psi_x + J \vec \Psi_x - T_s \Psi_s  - \vec 1 = 0\,,
 \\
 &
 (\vec \Psi_s)_\nu \in \{ 0, 1\}
\end{align}
where the nearest-neighbor sum is expressed by $J$ (zero diagonal and symmetric for non-directional graphs), the adjacency matrix for $G$.
The algorithm uses
\begin{equation}
    N_q = \overline{\overline{V}} + \sum_{v \in V} \log_2 \mathcal{N}(v)
\end{equation}
qubits to encode the vertices and to embed the slack variables.
Therefore, the (logical) qubit vector $\vec \Psi$ at worst scales with $n_V \log_2 n_V$ qubits for fully connected graphs.

The target QUBO in the notation of Eq.~\eqref{eq:matrix_form} reads
 {\small
  \begin{align}
   Q_{xx} & = \mathbbm{1} + p\left[J^T J + J^T + J - 2 \mathrm{diag}(|J|) - \mathbbm{1} \right] \,, \\
   Q_{sx} & = - p(\mathbbm{1}+J^T)T_s\,,                                                                     \\
   Q_{ss} & = p\left[{T_s}^T T_s + 2\mathrm{diag}(|T_s|)\right]\,,                                  \\
   C      & =  p \overline{\overline{V}}\,,
  \end{align}}
where $ |M| \equiv \sum_{\nu} M_{\nu \mu}$.

>>>>

The solution to the Minimum Dominating Set problem can be expressed as an integer optimization problem given by,
\begin{align}
 f(\vec x) = & \min\left(\sum_{v \in V} x_v\right),                    \\
\end{align}
subject to
\begin{align}
 & x_v + \sum_{j \in \mathcal{N}(v)} x_j \geq 1, \\
 & x_v \in \{0, 1\}
\end{align}
where $\mathcal{N}(v)$ represents the set of all direct neighbors of vertex $v$ and the dimension of the dependent variable $x$ is the number of vertices $n_V \equiv \overline{\overline{V}}$.
The problem minimizes the number of vertices in $D$ with a binary variable $\vec x = \vec \Psi_x$ encoded by single qubit, subject to the constraint that at least one vertex in $\mathcal{N}(\nu)$ is in $D$.
For each vertex in $V$ we introduce slack variables
\begin{equation}
    \vec s \in \left\{ \mathbb{Z}^{n_V} \, \middle| \, 0 \leq s_{v} \leq |\mathcal{N}(v)| \quad \forall v\in V \right\} \, ,
\end{equation}
which is related to the qubit vector $\vec \Psi_s$ by $\vec s = T_s \vec \Psi_s$.
The inequality constraint is encoded by
\begin{align}
 f(\vec \Psi_x)
 =
 & \min\left(\vec 1 \cdot \vec \Psi_x \right) \, ,
\end{align}
subject to
\begin{align}
 &
 (\mathbbm{1} + J)\vec \Psi_x - T_s \Psi_s  - \vec 1 = 0\,,
 \\
 &
 (\vec \Psi_s)_\nu \in \{ 0, 1\}
\end{align}
where the nearest-neighbor sum is expressed by the adjacency matrix $J$ (zero diagonal and symmetric for non-directional graphs).
The algorithm uses
\begin{equation}
    N_q = \overline{\overline{V}} + \sum_{v \in V} \log_2 \mathcal{N}(v)
\end{equation}
qubits to encode the vertices and slack variables before embedding.
Therefore, the (logical) qubit vector $\vec \Psi$ at worst scales with $n_V \log_2 n_V$ qubits for fully connected graphs.

The target QUBO in the notation of Eq.~\eqref{eq:matrix_form} reads
 {\small
  \begin{align}
   Q_{xx} & = \mathbbm{1} + p\left[J^T J + J^T + J - \mathrm{diag}\left(J^T \vec 1 + \vec 1^T J\right) - \mathbbm{1} \right] \,, \\
   Q_{sx} & = - p(\mathbbm{1}+J)^TT_s\,,                                                                     \\
   Q_{ss} & = p\left[{T_s}^T T_s + \mathrm{diag}\left(T_s^T \vec 1 + \vec 1^T  T_s\right)\right]\,,                                  \\
   C      & =  p \overline{\overline{V}}\,.
  \end{align}}
